\speculake: Hiding Malware in Speculative Execution

### 2018-08-08 paper/architecture.tex
Arch update and shortened to 15 pages

\section{Architecture}

\speculake malware is comprised of two independent programs: a payload program,
and a trigger program. Both are installed on the victim's computer (e.g. via
trojan, remote exploit, or phishing), and must run simultaneously on the same
physical CPU. We note that the constraint of running programs on the same CPU is
not a significant burden: \texttt{taskset} can be used to limit a process to a
core, or if not available, the attacker can simply run multiple copies of the
trigger or payload program to coerce the OS scheduler to assign a
payload/trigger pair to the same core. We note that this is not a strict
requirement for the attacker as most OS schedulers will use round robin
algorithms to assign tasks to processors---eventually the trigger and payload
will execute co-resident on a single core.

@@

@@

We found that the processor can identify some operations that have no effect on
output registers and allows them to use the entire size provided by the ROB, as
determined in Section~\ref{sssec:ROB}. This includes idiomatic no-ops and zero
idioms, which have no reliance on register values, for example the instruction
\texttt{xchg \%rax, \%rax} and \texttt{xor \%rax, \%rax}. However, if an
instruction is a potential data hazard, it must use an entry in the Physical 
Register File (PRF) until it reaches the correct stage in the execution pipeline. 
In this case, the number of instructions could be limited by the PRF
entries available.

%TODO: is this true? Or better explained by the micro-op story?

>>>>

\section{Architecture}
\speculake malware is comprised of two independent pieces: a payload program,
and a trigger. The payload, and some variants of the trigger, must be  installed
on the victim's computer(e.g. via trojan, remote exploit, or phishing). A running
payload performs innocuous measurements while waiting for the trigger to
become present.

If the trigger is a local program interacting with the payload
via the indirect branch predictor both programs must run on the same physical CPU. We 
note that this constraint is not a significant burden as most OS schedulers
will use round robin algorithms to assign tasks to processors---eventually 
the trigger and payload will execute co-resident on a single core.

@@

If the trigger is an input the program parses innocuous entries waiting 
for a series of crafted entries to train the direct branch predictor 
before a specially crafted trigger packet. The payload then forces the CPU
to speculate by stalling a bounds check. A local array is then speculatively 
overflowed redirecting program execution speculatively 
to an address within the payload program specified by the specially 
crafted trigger packet provided by the attacker.

@@

% We found that the processor can identify some operations that have no effect on
% output registers and allows them to use the entire size provided by the ROB, as
% determined in Section~\ref{sssec:ROB}. This includes idiomatic no-ops and zero
% idioms, which have no reliance on register values, for example the instruction
% \texttt{xchg \%rax, \%rax} and \texttt{xor \%rax, \%rax}. However, if an
% instruction is a potential data hazard, it must use an entry in the Physical 
% Register File (PRF) until it reaches the correct stage in the execution pipeline. 
% In this case, the number of instructions could be limited by the PRF
% entries available.
% 
% %TODO: is this true? Or better explained by the micro-op story?

### 2018-08-08 paper/discussion.tex
Arch update and shortened to 15 pages

%\paragraph{Red-Pills}
%Hardware minutia have also been used to identify and fingerprint execution
%environments---in malware this is known as a red-pill~\cite{red-pill} and is
%typically used by malware to detect when it is under emulation or a debugging
%environment~\cite{lindorfer2011detecting, balzarotti2010efficient,
%paleari2009fistful} attempting to circumvent malware detectors. To address
%red-pills, systems often employ measures to detect them.
\paragraph{Anti-Virus Detectors}
We verified that modern Anti-Virus technologies were unable to detect
and flag \speculake malware. We used ClamAV, BitDefender and rkhunter, which
mainly rely on signature- and string-based detection. BitDefender does feature
support for unpacking or extracting malware, though appears to simply try
unpacking using several known packers and encoding formats~\cite{BitDefender}.
Thus, it is not surprising that these tools cannot detect
\speculake.

@@

Alternatively an analyst may attempt to identify sections of the program or dead 
code that will be used to access the probe array and thereby find the speculative 
gadgets. However, identifying sections of memory that will access the probe 
array is equivalent to the "Must Alias" or "Points-To Problem" which has been 
proven undecidable without significant 
restrictions~\cite{ramalingam1994undecidability,landi1992undecidability}.

@@

\subsubsection{Alternative Applications}
We have proposed the \speculake model as a new malware threat, however, this
model of hidden computation could be applied more broadly. As of now there are
no ways of debugging speculative computation, meaning that the computation done
there is effectively isolated and difficult to extract by third-parties, even
with physical or root access to the machine it operates on. While not a true
trusted execution environment, it is possible that such an environment could be
used as a limited trusted environment. This may have potential use in digital
rights management (DRM), and whitebox cryptography applications.

>>>>

%\paragraph{Red-Pills}
%Hardware minutia have also been used to identify and fingerprint execution
%environments---in malware this is known as a red-pill~\cite{red-pill} and is
%typically used by malware to detect when it is under emulation or a debugging
%environment~\cite{lindorfer2011detecting, balzarotti2010efficient,
%paleari2009fistful} attempting to circumvent malware detectors. To address
%red-pills, systems often employ measures to detect them.
\paragraph{Anti-Virus Detectors}
We verified that modern Anti-Virus technologies were unable to detect
and flag \speculake malware. We used ClamAV, BitDefender and rkhunter, which
mainly rely on signature and string based detection. BitDefender does feature
support for unpacking or extracting malware, though appears to simply try
unpacking using several known packers and encoding formats~\cite{BitDefender}.
Thus, it is not surprising that these tools cannot detect
\speculake.

@@

Alternatively an analyst may attempt to identify sections of the program or dead 
code that will be used to access the probe array and thereby find the speculative 
gadgets. However, identifying sections of memory that will access the probe 
array is equivalent to the ``Must Alias'' or ``Points-To Problem'' which has been 
proven undecidable without significant 
restrictions~\cite{ramalingam1994undecidability,landi1992undecidability}.

@@

% \subsubsection{Alternative Applications}
% We have proposed the \speculake model as a new malware threat, however, this
% model of hidden computation could be applied more broadly. As of now there are
% no ways of debugging speculative computation, meaning that the computation done
% there is effectively isolated and difficult to extract by third-parties, even
% with physical or root access to the machine it operates on. While not a true
% trusted execution environment, it is possible that such an environment could be
% used as a limited trusted environment. This may have potential use in digital
% rights management (DRM), and whitebox cryptography applications.
%
